A Dearth of Young Stars and Planets in the Kepler Field

### 2024-04-26 paper/ms.tex
update ms

% <250 words
\begin{abstract}
  Recent studies of stars in open clusters have shown how the rotation
  rates and lithium content of FGK stars evolve during the first few
  billion years of their lives.
  %
  Here we apply this knowledge to planets and stars observed by the
  primary Kepler mission, and calculate the implied rotation and
  lithium-based ages.
  %
  Rotation periods are drawn from previous Kepler analyses; lithium is
  measured from new and archival HIRES spectra.
  %
  We report rotation-based ages for \nuniqstarsantosrotgyroappl\ stars
  and \nplwgyroagenograzing\ planets for which we believe
  gyrochronology is probably valid.
  %
  The independent lithium ages show X,Y,Z, though they are subject to
  their own systematic issues.
  %
  The relative statistical precisions from rotation, $\sigma_{t_{\rm
  gyro}} / t_{\rm gyro}$, are XX +YY -ZZ\%, which is X.X +Y.Y -Z.Z more
  statistically precise than isochronal ages for the same stars, and
  $\approx$X.X times less precise than cluster-based ages.
  %
  The results include \nplyounggyrotwosigma\ planets younger than
  1\,Gyr at 2$\sigma$ (\nplyounggyrothreesigma\ planets at 3$\sigma$).
  %
  While lithium provides only marginal added information for most of
  this sample, we report new robust ages from both rotation and
  lithium for 14 confirmed $<$500\,Myr planets, including five
  mini-Neptunes, five super-Earths, three Earths and sub-Earths, and a
  Neptune-sized planet.
  %
  In the broader Kepler target star sample, we find that the
  gyrochrone age distribution rises linearly from 0 to 3\,Gyr, with
  stars 2\,Gyr old being twice as common in the Kepler field as stars
  1\,Gyr old.
  %
  This parent stellar age distribution bounds the relative balance of
  young and old planets detectable not only by Kepler, but also by
  future wide-field transit surveys.
  %
\end{abstract}

@@

\begin{figure*}[!t]
	\begin{center}
		\subfloat{
			\includegraphics[width=0.47\textwidth]{st_params_M_G_vs_dr3_bp_rp.pdf}
		}
%	
%		\subfloat{
%			\includegraphics[width=0.4\textwidth]{st_params_M_G_vs_adopted_Teff.pdf}
%		}
%		\vspace{-0.5cm}
		\subfloat{
		\includegraphics[width=0.47\textwidth]{st_params_adopted_logg_vs_adopted_Teff.pdf}
		}
%		\subfloat{
%			\includegraphics[width=0.5\textwidth]{st_params_dr3_phot_g_mean_mag_vs_adopted_Teff.pdf}
%		}
	\end{center}
	\vspace{-0.5cm}
	\caption{
    {\bf The stellar samples.}  Our analysis focuses on stars observed
    by Kepler with previously reported rotation periods (dark gray
    points).  About half of these rotators are suitable for
    gyrochronology (black points), based on factors including their
    temperatures, surface gravities, (non-)binarity, and proximity to
    the main sequence.  The full list of criteria is given in
    Section~\ref{subsec:flags}.  A few percent of the remainder are
    either ``confirmed'' or ``candidate'' Kepler Objects of Interest
    that meet an additional set of quality criteria (blue points;
    Section~\ref{subsec:plflags}).  These are the objects for which
    our rotation-based ages are most reliable.  We also report
    lithium-based ages for a largely-overlapping setting of KOIs based
    on high-resolution spectroscopy.  Surface gravities and effective
    temperatures are primarily from
    \citet{Berger_2020a_catalog}.
	}
	\label{fig:stellarprops}
\end{figure*}

@@

\begin{figure*}[!t]
	\begin{center}
		\subfloat{
			\includegraphics[width=0.48\textwidth]{prot_teff_Santos19_Santos21_dquality.pdf}
		}
		%	
		%		\vspace{-0.5cm}
		\subfloat{
			\includegraphics[width=0.48\textwidth]{koi_mean_prot_teff_koi_X_S19S21dquality_keepgrazing.pdf}
		}
	\end{center}
	\vspace{-0.5cm}
	\caption{
		{\bf Rotation periods for Kepler stars (top) and planets
			(bottom)}.  The periods are from \citet{Santos_2019} and
		\citet{Santos_2021} (\citetalias{Santos_2019} and
		\citetalias{Santos_2021}).  The gray lines are ``mean fits'' from
		Figure~\ref{fig:models} to the slow rotation sequences of M67,
		Ruprecht~147, NGC~6811, Praesepe, and a trio of 120\,Myr open
		clusters.  The stellar sample shows only apparently single stars
		near the main sequence with $\log g$$>$4.2, RUWE$<$1.4,
		3800$<$$T_{\rm eff}$/K$<$6200, that are not flagged as eclipsing
		binaries.  The planet sample has the same stellar cuts, and shows
		confirmed and candidate planets from the cumulative KOI table,
		requiring them to have a MES$>$10.
	}
	\label{fig:prot_vs_teff}
\end{figure*}

>>>>

% <250 words
\begin{abstract}
  Recent studies of stars in open clusters have shown how the rotation
  rates and lithium content of FGK stars evolve during the first few
  billion years of their lives.
  %
  Here we apply this knowledge to stars observed by the
  primary Kepler mission, and calculate the implied rotation and
  lithium-based ages.
  %
  Rotation periods are drawn from previous Kepler analyses; lithium is
  measured from new and archival HIRES spectra.
  %
  We report rotation-based ages for \nuniqstarsantosrotgyroappl\ stars
  and \nplwgyroagenograzing\ planets for which we believe
  gyrochronology is valid.
  %
  The independent lithium ages show X,Y,Z, though they are subject to
  their own systematic issues.
  %
  The relative statistical precisions from rotation, $\sigma_{t_{\rm
  gyro}} / t_{\rm gyro}$, are XX +YY -ZZ\%, which is X.X +Y.Y -Z.Z more
  statistically precise than isochronal ages for the same stars, and
  $\approx$X.X times less precise than cluster-based ages.
  %
  The results include \nplyounggyrotwosigma\ planets younger than
  1\,Gyr at 2$\sigma$ (\nplyounggyrothreesigma\ planets at 3$\sigma$).
  %
  While lithium provides only marginal added information for most of
  this sample, we report new robust ages from both rotation and
  lithium for 14 confirmed $<$500\,Myr planets, including five
  mini-Neptunes, five super-Earths, three Earths and sub-Earths, and a
  Neptune-sized planet.
  %
  In the broader Kepler target star sample, we find that the
  gyrochrone age distribution rises linearly from 0 to 3\,Gyr, with
  stars 2\,Gyr old being twice as common in the Kepler field as stars
  1\,Gyr old.
  %
  This parent stellar age distribution bounds the relative balance of
  young and old planets detectable not only by Kepler, but also by
  future wide-field transit surveys.
  %
\end{abstract}

@@

\begin{figure*}[!t]
	\begin{center}
		\subfloat{
			\includegraphics[width=0.47\textwidth]{st_params_M_G_vs_dr3_bp_rp.pdf}
		}
%	
%		\subfloat{
%			\includegraphics[width=0.4\textwidth]{st_params_M_G_vs_adopted_Teff.pdf}
%		}
%		\vspace{-0.5cm}
		\subfloat{
		\includegraphics[width=0.47\textwidth]{st_params_adopted_logg_vs_adopted_Teff.pdf}
		}
%		\subfloat{
%			\includegraphics[width=0.5\textwidth]{st_params_dr3_phot_g_mean_mag_vs_adopted_Teff.pdf}
%		}
	\end{center}
	\vspace{-0.5cm}
	\caption{
    {\bf The stars.}  Our analysis focuses on stars observed
    by Kepler with previously reported rotation periods (dark gray
    points).  About half of these rotators are suitable for
    gyrochronology (black points), based on factors including their
    temperatures, surface gravities, (non-)binarity, and proximity to
    the main sequence.  The full list of criteria is given in
    Section~\ref{subsec:flags}.  A few percent of the remainder are
    either ``confirmed'' or ``candidate'' Kepler Objects of Interest
    that meet an additional set of planetary quality criteria (blue points;
    Section~\ref{subsec:plflags}).  
%	We independently derive
%    lithium-based ages for a largely-overlapping setting of KOIs based
%    on high-resolution spectroscopy.
    Surface gravities and effective
    temperatures are primarily from
    \citet{Berger_2020a_catalog}.
	}
	\label{fig:stellarprops}
\end{figure*}

@@

\begin{figure*}[!t]
	\begin{center}
		\subfloat{
			\includegraphics[width=0.48\textwidth]{prot_teff_Santos19_Santos21_dquality.pdf}
		}
		%	
		%		\vspace{-0.5cm}
		\subfloat{
			\includegraphics[width=0.48\textwidth]{koi_mean_prot_teff_koi_X_S19S21dquality_keepgrazing.pdf}
		}
	\end{center}
	\vspace{-0.5cm}
	\caption{
		{\bf Rotation periods for Kepler stars (left) and planets
			(right)}.  The periods are from \citet{Santos_2019} and
		\citet{Santos_2021} (\citetalias{Santos_2019} and
		\citetalias{Santos_2021}).  The gray lines are ``mean fits'' from
		Figure~\ref{fig:models} to the slow rotation sequences of M67,
		Ruprecht~147, NGC~6811, Praesepe, and a trio of 120\,Myr open
		clusters.  The stellar sample shows only apparently single stars
		near the main sequence with $\log g$$>$4.2, RUWE$<$1.4,
		3800$<$$T_{\rm eff}$/K$<$6200, that are not flagged as eclipsing
		binaries.  The planet sample has the same stellar cuts, and shows
		confirmed and candidate planets from the cumulative KOI table,
		requiring them to have a MES$>$10.
	}
	\label{fig:prot_vs_teff}
\end{figure*}
